Data Predictive Control with Random Forests:\\Theory and Applications\tnoteref{mytitlenote}

### 2017-09-26 paper/abstract.tex
completed abstract, section 5 and conclusion

Model-based control techniques have been widely used over the past years in control systems to improve system performance of simple controllers, such as PID, bang-bang, etc. Among these, one of the most popular is Model Predictive Control (MPC), due to its capacity to optimally control a system, based on the capability to predict its behavior over a future horizon. A key factor prohibiting the widespread adoption of MPC for complex systems is related to the difficulties (cost, time, and effort) associated with the model identification. Data-driven models have been intensively studied in the last year to overcome this problem. However, closed-loop control systems based on such models have not received much attention. We introduce a novel idea for predictive control based on system data, Data Predictive Control (DPC), leveraging machine learning algorithms, in particular regression trees and random forests. We show, using a bilinear model, that DPC has comparable performance with MPC, and apply it to the Demand Response problem of a building and to find the optimal ON/OFF scheduling of the heating system of a real house in order to minimize the power consumption while guaranteeing thermal comfort for occupants.

>>>>

Model-based control techniques have been widely used over the past years in control systems to improve system performance of simple controllers, such as PID, bang-bang, etc. Among these, one of the most popular is Model Predictive Control (MPC), due to its capacity to optimally control a system, based on the capability to predict its behavior over a future horizon. A key factor prohibiting the widespread adoption of MPC for complex systems is related to the difficulties (cost, time, and effort) associated with the model identification. Data-driven models have been intensively studied in the last year to overcome this problem. However, closed-loop control systems based on such models have not received much attention. We introduce a novel idea for predictive control based on system data, Data Predictive Control (DPC), leveraging machine learning algorithms, in particular regression trees and random forests. We show, using a bilinear model, that DPC has comparable performance with MPC. We apply it to the Demand Response problem of a large scale multi-story EnergyPlus building model, and show that DPC curtails the desired power usage with high confidence. Finally we apply it to find the optimal ON/OFF scheduling of the heating system of a real house in order to minimize the power consumption while guaranteeing thermal comfort for occupants. Results show that, compared to a bang-bang controller, with DPC we obtain an energy saving in terms of $kWh$ that goes from $25,4\%$ to $49.2\%$.

### 2017-09-26 paper/casestudyReal.tex
completed abstract, section 5 and conclusion

\begin{figure}[h!]
	\begin{center}
		\includegraphics[width=26pc]{figures/Schema_di_impianto.eps}
		\caption{Technological plant scheme of the use case.}
		\captionsetup{justification=centering}
		\label{F:housePlantScheme}
	\end{center}
\end{figure}
For a thorough knowledge of the building thermal behavior, an in-situ analysis was carried out. Temperature measuring devices, completely self-produced at the “G. Parolini Lab” of the University of L'Aquila \cite{Pantoli2017}, mainly based on an $ATmega 2560$ microcontroller and $DS18B20$ temperature probes (temperature range from $-55.0\degree C$ to $125.0\degree C$ with an accuracy of $\pm 0.5\degree C$), were employed to acquire the ambient temperatures of the four differently oriented rooms of the house as shown in Figure \ref{F:houseFloors}. The probes positions were chosen to optimize the data acquisition and to minimize the discomfort of the occupants in the house. Furthermore, as can be seen in Figures \ref{F:housePlantScheme} and \ref{F:houseFloors} (COMMENT: NOT CLEAR WHERE THE SENSORS ARE IN THE FIGURES. MORE DETAILS NEEDED), a commercial heat meter (temperature range from $10.0\degree C$ to $90.0\degree C$ with an accuracy of $\pm 0.05\degree C$) was installed downstream of the biomass boiler to measure the produced thermal energy. All the measuring devices have been set with a data acquisition rate equal to $10$ minutes, from March $11^{th}$ $2016$ to May $15^{th}$ $2016$.
\begin{figure}[h!]
	\begin{center}
		\subfigure[Ground floor.]{
			\label{F:houseGroundFloor}
			\centering
			\includegraphics[width=26pc]{figures/Arch_PT_e_sonde.eps}
		}
		\subfigure[Attic.]{
			\label{F:houseAttic}
			\centering
			\hspace{-1.6cm}
			\includegraphics[width=22.8pc]{figures/Arch_P1.eps}
		}
	\end{center}
	\caption{Layout of the house and probes placement. Legend: red circle for ambient temperature; orange circle for heat meter.}
	\captionsetup{justification=centering}
	\label{F:houseFloors}
\end{figure}

@@

\subsection{EnergyPlus model}\label{SS:energyPlusmodel}
To assess the energy performance of the use case's heating system, the EnergyPlus engine coupled with DesignBuilder modeling environment was employed. Based on the weather data provided by CETEMPS – Centre of Excellence \cite{CETEMPS}, shown in Figure \ref{F:houseExternalWeather}, the performed dynamic simulation was based on a weather file specifically created for L'Aquila, that according to the K\"{o}ppen-Geiger climate classification, has a warm temperate climate (Csa, Cfa and Csb) \cite{Peel2007}. (COMMENT: TO BE BETTER WRITTEN)

\begin{figure}[h!]
	\begin{center}
		\includegraphics[width=26pc]{figures/dati_climatici_rev01.eps}
		\caption{Weather data of L'Aquila.}
		\captionsetup{justification=centering}
		\label{F:houseExternalWeather}
	\end{center}
\end{figure}

@@

\begin{figure}[h!]
	\begin{center}
		\includegraphics[width=28pc]{figures/Scheduling_rev02.eps}
		\caption{Actual scheduling of the heating system, where white color means switched OFF and green color switched ON, and ($T_{ext,ave}$) is the external average temperature. From midnight to $8am$ the heating system is always switched off.)}
		\captionsetup{justification=centering}
		\label{F:houseActualScheduling}
	\end{center}
\end{figure}
Following the hourly calibration proposed by the M\&V guidelines of ASHRAE \cite{USDOE}, a simulated model is calibrated when the mean bias error (MBE) and the coefficient of variation of the root mean square error [CV(RMSE)] are less than acceptable tolerances, respectively equal to $\pm 10.0\%$ and $30.0\%$. MBE and CV(RMSE) have been calculated by using Equations \eqref{Eq:mbe} and \eqref{Eq:cvrmse}.

@@

We use random forest model $\tP^r$ also to estimate the energy consumption of the house. In particular we estimated it for the same period analyzed in Section \ref{SS:energyPlusmodel} and obtained the results in Figure \ref{F:Energy_testing}, with a MBE equal to $13.0\%$ and a CV(RMSE) equal to $13.7\%$.

\begin{figure}[h!]
	\begin{center}
		\includegraphics[width=26pc]{figures/Energy_test.eps}
		\caption{Energy consumption accuracy.}
		\captionsetup{justification=centering}
		\label{F:Energy_testing}
	\end{center}
\end{figure}

As we can see, with random forest model we get a worse MBE and CV(RMSE) for energy consumption, although the maximum error is very similar. This is because the EnergyPlus model, due to its internal dynamics based on the house structure, is able to compensate the error over time. On the other hand our curve perfectly follows the time behavior of the real case. For this reason, in Section \ref{SS:simulationResults}, we will use both models in order to compare the quality of the DPC with respect to a classical bang-bang controller.

@@

We ran DPC with $3$ different sets of parameters $Q$, $\lambda_{min}$ and $\lambda_{max}$. Each set allows a different level of temperature bounds violation. In particular we considered a small, a medium and a large violation configuration. The simulation period was from May $1^{st}$ $2016$ at $00am$ to May $5^{th}$ $2016$ at $1pm$.
\paragraph{Result 1} For the sake of clarity, we report in Figure \ref{F:comparison_small} only the comparison for temperature and input schedule obtained allowing small bound violations in DPC. We can see that the temperature controlled with DPC does not violate the bounds and if it does then the violation is around $0.1\,\degree C$. Bang-bang control also presents small bounds violations due to its working principle. We can see how the DPC control law requires the heating system to be ON for less time than the bang-bang one to keep the temperature in the comfort range. We will see in Figure \ref{F:comparison_all_energy} that this translates in energy saving.

@@

\paragraph{Result 3} In Figure \ref{F:comparison_all_energy}, using the energy model derived from random forest power model $\tP^r$, we show how DPC outperforms the bang-bang controller also in terms of energy consumption and how the bounds violations allows us to save more energy. We can see that if we want to keep the temperature within a comfort range, only allowing small violations, the use of DPC produce an energy saving of $44\, kWh$, that correspond to the $33.0\%$ over a period of $4$ days and a half. Instead if we allow a large violation over the same period, the energy saving is of $78\, kWh$, that correspond to the $59.0\%$. If we consider these results over a monthly period, i.e. multiplying them for $6$, we get a monthly energy saving that goes from $264\,kWh$ to $468\,kWh$, with different comfort constraint configurations. Considering that the equivalent price of $1\,kWh$ of vegetable biomass is $TOT$\euro{} (TULLIO MAY YOU TELL ME ABOUT THIS PART?), we have a monthly saving of $TOT$\euro{}.

\begin{figure}[h!]
	\begin{center}
		\includegraphics[width=26pc]{figures/Energy_all.eps}
	\end{center}
	\caption{Comparison of DPC and bang-bang control performance with different violations using random forest model.}
	\label{F:comparison_all_energy}
\end{figure}
An alternative comparison can be done considering the EnergyPlus model of the house for energy consumption described in Section \ref{SS:energyPlusmodel}. Results are shown in Figure \ref{F:comparison_all_energy_E+}.
\begin{figure}[h!]
	\begin{center}
		\includegraphics[width=26pc]{figures/Energy_all.eps}
	\end{center}
	\caption{SUBSTITUTE WITH TULLIO'S RESULTS Comparison of DPC and bang-bang control performance with different violations using EnergyPlus model.}
	\label{F:comparison_all_energy_E+}
\end{figure}
\textbf{TO BE ADAPTED WITH TULLIO'S RESULTS We can see that, over the same period of $4$ days and a half, using DPC we saved $TOT\, kWh$ allowing small violations, that correspond to a $TOT\%$ of energy saving, and $TOT\, kWh$ allowing large violations, that correspond to a $TOT\%$ of energy saving.}

>>>>

\begin{figure}[h!]
	\begin{center}
		\includegraphics[width=26pc]{figures/Schema_di_impianto.eps}
		\caption{Technological plant scheme of the use case.}
		\captionsetup{justification=centering}
		\label{F:housePlantScheme}
	\end{center}
\end{figure}
For a thorough knowledge of the building thermal behavior, an in-situ analysis was carried out. Temperature measuring devices, completely self-produced at the “G. Parolini Lab” of the University of L'Aquila \cite{Pantoli2017}, mainly based on an $ATmega 2560$ microcontroller and $DS18B20$ temperature probes (temperature range from $-55.0\degree C$ to $125.0\degree C$ with an accuracy of $\pm 0.5\degree C$), were employed to acquire the ambient temperatures of the four differently oriented rooms of the house as shown in Figure \ref{F:houseFloors}. The probes positions were chosen to optimize the data acquisition and to minimize the discomfort of the occupants in the house. Furthermore, as can be seen in Figures \ref{F:housePlantScheme} and \ref{F:houseFloors}, a commercial heat meter (temperature range from $10.0\degree C$ to $90.0\degree C$ with an accuracy of $\pm 0.05\degree C$) was installed downstream of the biomass boiler to measure the produced thermal energy. The turbine flowmeter was installed on the return pipe, while the two thermocouples were placed inside the delivery and return pipes, respectively. All the measuring devices have been set with a data acquisition rate equal to $10$ minutes, from March $11^{th}$ $2016$ to May $15^{th}$ $2016$.
\begin{figure}[h!]
	\begin{center}
		\subfigure[Ground floor.]{
			\label{F:houseGroundFloor}
			\centering
			\includegraphics[width=26pc]{figures/Arch_PT_e_sonde.eps}
		}
		\subfigure[Attic.]{
			\label{F:houseAttic}
			\centering
			\hspace{-1.6cm}
			\includegraphics[width=22.8pc]{figures/Arch_P1.eps}
		}
	\end{center}
	\caption{Layout of the house and probes placement. Legend: red circle for ambient temperature; orange circle for heat meter.}
	\captionsetup{justification=centering}
	\label{F:houseFloors}
\end{figure}

@@

\subsection{EnergyPlus model}\label{SS:energyPlusmodel}
In this section we create an EnergyPlus model of the thermal energy consumption of the house that will be used in Section \ref{SS:simulationResults} to compare the quality of the DPC with respect to a classical bang-bang controller.

To assess energy performance of the use case heating system, the EnergyPlus engine together with DesignBuilder modeling environment has been employed. The model has been created according to the K\"{o}ppen-Geiger climate classification that, based weather data of L'Aquila shown in Figure \ref{F:houseExternalWeather}, provided by the CETEMPS – Centre of Excellence \cite{CETEMPS}, classifies Italy in Csa, Cfa and Csb climate zones, with a warm temperate climate [4], .

\begin{figure}[h!]
	\begin{center}
		\includegraphics[width=30pc]{figures/dati_climatici_rev01.eps}
		\caption{Weather data of L'Aquila.}
		\captionsetup{justification=centering}
		\label{F:houseExternalWeather}
	\end{center}
\end{figure}

@@

\begin{figure}[h!]
	\begin{center}
		\includegraphics[width=28pc]{figures/Scheduling_rev02.eps}
		\caption{Actual scheduling of the heating system, where white color means switched OFF and green color switched ON, and ($T_{ext,ave}$) is the external average temperature. From midnight to $8am$ the heating system is always switched off.}
		\captionsetup{justification=centering}
		\label{F:houseActualScheduling}
	\end{center}
\end{figure}
Following the hourly calibration proposed by the M\&V guidelines of ASHRAE \cite{USDOE}, also applied in \cite{Mustafarai2014,Raftery2011}, a simulated model is calibrated when the mean bias error (MBE) and the coefficient of variation of the root mean square error [CV(RMSE)] are less than acceptable tolerances, respectively equal to $\pm 10.0\%$ and $30.0\%$. MBE and CV(RMSE) have been calculated by using Equations \eqref{Eq:mbe} and \eqref{Eq:cvrmse}.

@@

%We use random forest model $\tP^r$ also to estimate the energy consumption of the house. In particular we estimated it for the same period analyzed in Section \ref{SS:energyPlusmodel} and obtained the results in Figure \ref{F:Energy_testing}, with a MBE equal to $13.0\%$ and a CV(RMSE) equal to $13.7\%$.
%
%\begin{figure}[h!]
%	\begin{center}
%		\includegraphics[width=26pc]{figures/Energy_test.eps}
%		\caption{Energy consumption accuracy.}
%		\captionsetup{justification=centering}
%		\label{F:Energy_testing}
%	\end{center}
%\end{figure}
%
%As we can see, with random forest model we get a worse MBE and CV(RMSE) for energy consumption, although the maximum error is very similar. This is because the EnergyPlus model, due to its internal dynamics based on the house structure, is able to compensate the error over time. On the other hand our curve perfectly follows the time behavior of the real case. For this reason, in Section \ref{SS:simulationResults}, we will use both models in order to compare the quality of the DPC with respect to a classical bang-bang controller.

@@

We ran DPC with $3$ different sets of parameters $Q$, $\lambda_{min}$ and $\lambda_{max}$. Each set allows a different level of temperature bounds violation. In particular we considered a small, a medium and a large violation configuration.
\paragraph{Result 1} For the sake of clarity, we report in Figure \ref{F:comparison_small} only the comparison for temperature and input schedule obtained allowing small bound violations in DPC.  For the same reason the simulation period is restricted to $4$ days and a half, from May $1^{st}$ $2016$ at $00am$ to May $5^{th}$ $2016$ at $1pm$, in order to make the plots more clear. We can see that the temperature controlled with DPC does not violate the bounds and if it does then the violation is around $0.1\,\degree C$. Bang-bang control also presents small bounds violations due to its working principle. We can see how the DPC control law requires the heating system to be ON for less time than the bang-bang one to keep the temperature in the comfort range. We will see in Figure \ref{F:comparison_all_energy_E+} that this translates in energy saving.

@@

\paragraph{Result 3} 
%In Figure \ref{F:comparison_all_energy_E+}, using the energy model derived from random forest power model $\tP^r$, we show how DPC outperforms the bang-bang controller also in terms of energy consumption and how the bounds violations allow us to save more energy. We can see that if we want to keep the temperature within a comfort range, only allowing small violations, the use of DPC produce an energy saving of $44\, kWh$, that correspond to the $33.0\%$ over a period of $4$ days and a half. Instead if we allow a large violation over the same period, the energy saving is of $78\, kWh$, that correspond to the $59.0\%$. If we consider these results over a monthly period, i.e. multiplying them for $6$, we get a monthly energy saving that goes from $264\,kWh$ to $468\,kWh$, with different comfort constraint configurations. Considering that the equivalent price of $1\,kWh$ of vegetable biomass is $TOT$\euro{} (TULLIO MAY YOU TELL ME ABOUT THIS PART?), we have a monthly saving of $TOT$\euro{}.

%\begin{figure}[h!]
%	\begin{center}
%		\includegraphics[width=26pc]{figures/Energy_all.eps}
%	\end{center}
%	\caption{Comparison of DPC and bang-bang control performance with different violations using random forest model.}
%	\label{F:comparison_all_energy}
%\end{figure}
%An alternative comparison can be done considering the EnergyPlus model of the house for energy consumption described in Section \ref{SS:energyPlusmodel}. Results are shown in Figure \ref{F:comparison_all_energy_E+}.
In Figure \ref{F:comparison_all_energy_E+}, using the thermal energy consumption model derived in Section \ref{SS:energyPlusmodel} using EnergyPlus, we show how DPC outperforms the bang-bang controller also in terms of energy consumption and how the bounds violations allow us to save more energy. In this case, since the plot is clear and we are interested in showing energy saving on a long period, we ran the simulations over the testing data, i.e. $15$ days of May, from May $1^{st}$ $2016$ to May $15^{th}$ $2016$. We can see that if we want to keep the temperature within a comfort range, only allowing small violations, the use of DPC produce an energy saving of $45\, kWh$, that correspond to the $25.4\%$ over a period of $15$ days. Instead if we allow a large violation over the same period, the energy saving is of $87\, kWh$, that correspond to the $49.2\%$.
\begin{figure}[h!]
	\begin{center}
		\includegraphics[width=26pc]{figures/Energy_all_EnergyPlus.eps}
	\end{center}
	\caption{Comparison of DPC and bang-bang controller performance in terms of thermal energy saving with different violations using EnergyPlus model.}
	\label{F:comparison_all_energy_E+}
\end{figure}

### 2017-09-26 paper/conclusion.tex
completed abstract, section 5 and conclusion

We present two algorithms based on trees and random forests for receding horizon control with data-driven models. 
We compare the performance of our Data Predictive Control to MPC on a multivariable bilinear building model. We establish that DPC with random forests shows a remarkable similarity to MPC in the optimal control strategies explaining 70\% variance. On the other hand, DPC with regression trees suffers from practical limitations due to model overfitting.
We further apply DPC with random forests to a large scale 6 story EnergyPlus model with 22 zones for which the traditional model-based control is largely unsuitable due to complex dynamics and the cost of model identification. We show that DPC, relying only on the sensor data, can provide significant energy savings while maintaining thermal comfort. Our results demonstrate that even for such complex system, DPC tracks a reference signal with a mean error of 3\%.

>>>>

\section{CONCLUSION}
\label{S:conclusion}

We present two algorithms based on trees and random forests for receding horizon control with data-driven models.

We compare the performance of our Data Predictive Control to MPC on a multivariable bilinear building model. We establish that DPC with random forests shows a remarkable similarity to MPC in the optimal control strategies explaining 70\% variance. On the other hand, DPC with regression trees suffers from practical limitations due to model overfitting.

We further apply DPC with random forests to a large scale 6 story EnergyPlus model with 22 zones for which the traditional model-based control is largely unsuitable due to complex dynamics and the cost of model identification. We show that DPC, relying only on the sensor data, can provide significant energy savings while maintaining thermal comfort. Our results demonstrate that even for such complex system, DPC tracks a reference signal with a mean error of 3\%.

We finally apply DPC using data from a real house and compare it with the classical bang-bang controller widely used for temperature control in houses. We implemented DPC to guarantee energy saving while guaranteeing thermal comfort for the occupants. We obtained, with respect to a bang-bang controller, an energy saving that goes from the $25.4\%$, when we want the temperature to be strictly in a comfort range, to the $49.2\%$, if we allow little violations of such a range.

### 2017-09-26 paper/dpc.tex
completed abstract, section 5 and conclusion

Consider a black-box model given by $x_{k+1}=f(x_k,u_k,d_k)$, where $x,u,d$ represent states, inputs and disturbances, respectively. Depending upon the learning algorithm, $f$ is typically nonlinear, nonconvex and sometimes nondifferentiable (as is the case with regression trees and random forests) with no closed-form expression. 
Such functional representations learned through black-box modeling may not be directly suitable for control and optimization as the optimization problem can be computationally intractable, or due to nondifferentiabilities we may have to settle with a sub-optimal solution using evolutionary algorithms \cite{Kusiak2009}.
These problems can be eliminated by decomposing 
\begin{align}
f(x_k,u_k,d_k)= g(d_k,x_k,h(u_k)),
\label{E:sepvars}
\end{align}
where both $g$ and $h$ are learned using the data, and $h(u_k)$ is convex and differentiable, and thus suitable for optimization. DPC uses this functional decomposition or \textit{separation of variables} to overcome the aforementioned challenges with black-box optimization.

@@

\subsection{DPC-En: DPC with Ensemble Methods}
\label{SS:dpcrf}
Regression trees obtain good predictive accuracy in many domains. However, the models used in their leaves have some limitations regarding the kind of functions they are able to approximate.
The problem with trees is their high variance and that they can overfit the data easily.
A small change $\delta$ in the data can result in a different series of splits and thus violate the acceptable accuracy $\eps$, i.e. $\exists \  \hat{\tX}^d \ | \ ||\tX^d-\hat{\tX}^d|| < \delta$ \& $||\tY-\tY_{\mathrm{true}}||>\eps$.
This is the price to be paid for estimating a tree-based structure from the data.
%The main reason behind this is the hierarchical nature of the process: the effect of an error in the top split is propagated down to all of the splits below it.
%While pruning and cross validation can help reduce the over fitting partially, we use ensemble methods for growing more accurate trees.

@@

The goal with DPC-En is to replace each tree in Alg.~\ref{A:dpcrt} by a forest
\begin{gather}
\label{E:model_forest}
\tY_{\mathrm{k+j|k}} = \mathit{f}_{\mathrm{forest}} \left( \tX^d_{\mathrm{k+j-N|k}} ,\dots,\tX^d_{\mathrm{k+j-1|k}}  \right), \\ \nonumber
\mathsf{X}^d_{\mathrm{k+j-l|k}} \in \R^p \  \ \forall \ l,j=1,\dots,N,
\end{gather}
which, again, is trained only on $\mathsf{X}^d$, but $\tX^d \in \R^p \subset \R^d$ for each tree,  and then fit a linear regression model using $\mathsf{X}^c$ in every leaf of every tree. We build $N$ such forests for $N$ prediction steps such that the leaf $R_i$ of forest $\mathcal{R}_j$ uses a linear model
\begin{gather}
\label{E:model_leaf_forest}
\tY_{\mathrm{k+j|k}} =  \Theta^T_{ij} [1,\tX^c_{\mathrm{k|k}},\dots,\tX^c_{\mathrm{k+j-1|k}} ]^T, \\ \nonumber
\mathsf{X}^c_{\mathrm{k+j-l|k}} \in \R^c \  \ \forall \ l,j=1,\dots,N.
\end{gather}
Here $(\mathsf{X}^c,\mathsf{Y})$ correspond to the in-bag samples for the trees.

>>>>

Consider a black-box model given by $x_{k+1}=f(x_k,u_k,d_k)$, where $x,u,d$ represent states, inputs and disturbances, respectively. Depending upon the learning algorithm, $f$ is typically nonlinear, nonconvex and sometimes nondifferentiable (as is the case with regression trees and random forests) with no closed-form expression. 
Such functional representations learned through black-box modeling may not be directly suitable for control and optimization as the optimization problem can be computationally intractable, or due to nondifferentiabilities we may have to settle with a sub-optimal solution using evolutionary algorithms \cite{Kusiak2009}.
These problems can be eliminated by decomposing 
\begin{align}
f(x_k,u_k,d_k)= g(x_k,h(u_k),d_k),
\label{E:sepvars}
\end{align}
where both $g$ and $h$ are learned using the data, and $h(u_k)$ is convex and differentiable, and thus suitable for optimization. DPC uses this functional decomposition or \textit{separation of variables} to overcome the aforementioned challenges with black-box optimization.

@@

\subsection{DPC-En: DPC with Ensemble Methods}
\label{SS:dpcrf}
Regression trees obtain good predictive accuracy in many domains. However, the models used in their leaves have some limitations regarding the kind of functions they are able to approximate.
The problem with trees is their high variance and that they can overfit the data easily.
A small change $\delta$ in the data can result in a different series of splits and thus violate the acceptable accuracy $\eps$.
This is the price to be paid for estimating a tree-based structure from the data.
%The main reason behind this is the hierarchical nature of the process: the effect of an error in the top split is propagated down to all of the splits below it.
%While pruning and cross validation can help reduce the over fitting partially, we use ensemble methods for growing more accurate trees.

@@

The goal with DPC-En is to replace each tree in Alg.~\ref{A:dpcrt} by a forest
\begin{gather}
\label{E:model_forest}
\tY_{\mathrm{k+j|k}} = \mathit{f}_{\mathrm{forest}} \left( \tX^d_{\mathrm{k+j-N|k}} ,\dots,\tX^d_{\mathrm{k+j-1|k}}  \right), \\ \nonumber
\mathsf{X}^d_{\mathrm{k+j-l|k}} \in \R^p \  \ \forall \ l,j=1,\dots,N,
\end{gather}
which, again, is trained only on $\mathsf{X}^d$, with $\tX^d \in \R^p \subset \R^d$ for each tree,  and then fit a linear regression model using $\mathsf{X}^c$ in every leaf of every tree. We build $N$ such forests for $N$ prediction steps such that the leaf $R_i$ of forest $\mathcal{R}_j$ uses a linear model
\begin{gather}
\label{E:model_leaf_forest}
\tY_{\mathrm{k+j|k}} =  \Theta^T_{ij} [1,\tX^c_{\mathrm{k|k}},\dots,\tX^c_{\mathrm{k+j-1|k}} ]^T, \\ \nonumber
\mathsf{X}^c_{\mathrm{k+j-l|k}} \in \R^c \  \ \forall \ l,j=1,\dots,N.
\end{gather}
Here $(\mathsf{X}^c,\mathsf{Y})$ correspond to the in-bag samples for the trees.

### 2017-09-26 paper/intro.tex
completed section 5 and fixed figures problems
completed abstract, section 5 and conclusion

In our previous work~\cite{JainACC2017,Jain2016}, we introduced the concept of DPC for receding horizon control.
This work has the following contributions.
(1) We first formally present two underlying algorithms: (i) DPC with regression trees, and (ii) DPC with random forests, which also ensure recursive feasibility in receding horizon control.
(2) Using a bilinear building model whose parameters were identified using experiments on a building in Switzerland, we demonstrate the strength of DPC for receding horizon control via one-to-one comparison against a benchmark MPC controller. We show DPC captures 70\% variance in MPC and offers a comparable performance.
(3) We present a practical application of DPC for Demand Response, where we apply DPC to a 6 story 22 zone building model in EnergyPlus \cite{Crawley2001} for which model-based control is not economical and practical due to extreme complexity. We show scalability and efficiency of DPC in providing financial incentives to the end-customers bypassing the need for high fidelity models. We observe that DPC provides the desired power reduction with an average error of 3\%.

>>>>

In our previous work~\cite{Behl2016,Jain2016,JainACC2017,JainCDC2017}, we introduced the concept of DPC for receding horizon control.
This work has the following contributions.
(1) We first formally present two underlying algorithms: (i) DPC with regression trees, and (ii) DPC with random forests.
(2) Using a bilinear building model whose parameters were identified using experiments on a building in Switzerland, we demonstrate the strength of DPC for receding horizon control via one-to-one comparison against a benchmark MPC controller. We show DPC captures 70\% variance in MPC and offers a comparable performance.
(3) We present a practical application of DPC for Demand Response, where we apply DPC to a 6 story 22 zone building model in EnergyPlus \cite{Crawley2001} for which model-based control is not economical and practical due to extreme complexity. We show scalability and efficiency of DPC in providing financial incentives to the end-customers bypassing the need for high fidelity models. We observe that DPC provides the desired power reduction with an average error of 3\%.
